Eliminating prior-bias from sparse-projection tomographic reconstructions

### 2020-07-18 paper/tci.tex
Minor changes in the intro

In conventional data acquisition techniques, the tomographic
measurements $\boldsymbol{y}$ are acquired by sampling the physical
object $\boldsymbol{x}$ uniformly with a substantial number of views,
ideally above the Nyquist rate. In such a case when there are
sufficient measurements, reconstruction using the conventional
filtered backprojection (FBP) suffices, as seen in
Fig.~\ref{fig:story} (first column, 600 views). The figure shows the
ground truth image (260$\times$260) of naturally growing sprouts at the top
left (the details of the dataset is postponed to
Section~\ref{sec:results_uniform_prior}).

@@

 \begin{figure}[t]
\centering
	\includegraphics[width=0.45\textwidth]{../images/story/story.png}
        \caption{Illustration of the use of various priors as the
          number of views reduces. For extremely large number of views
          (600 views), the FBP reconstruction (first column) is of
          very good quality (Structural Similarity Index Metric,
          SSIM=0.92). As the number of views become limited (100
          views), the reconstruction using sparsity prior (middle
          column) is better both visually and quantitatively than FBP
          (SSIM of 0.90 vs 0.83). If the number of views is
          drastically reduced (30 views), the presence of both
          sparsity prior and object prior improves the reconstruction
          (SSIM = 0.85) when compared to presence of only sparsity
          prior (SSIM = 0.82) or no prior (SSIM = 0.69). In all cases,
          the ground truth (260x260) appears in the top left.}
 \label{fig:story}
 \end{figure}

 When the number of views is significantly reduced, the sparsity prior
 alone is not sufficient.  In such cases additional information
 specific to the object being scanned (i.e., object prior) is useful
 in further improving the reconstruction.  While this has been done in
 the context of dynamic CT-scans (as discussed in
 Section~\ref{sec:related}), in this paper we use a set $S$ of
 previous scans of the same object that is more appropriate in a
 longitudinal context. Specifically, longitudinal studies here refers
 to the acquisition of sequential CT scans of the same subject to
 track time-evolving changes within the subject's interior. As seen in
 Fig.~\ref{fig:story} (last column, 30 views) our method uses the
 prior scans \emph{uniformly} by creating an eigenspace representation
 (Section~\ref{sec:method_uniform_prior}).

 However, with the use of such an object prior, a new challenge
 emerges. The prior set S may potentially overwhelm the necessary
 details, or since several prior scans are available, finding the
 right prior may be an issue. We therefore seek an algorithm that
 estimates the location and magnitude of new changes in the (unknown)
 test. As we show in Section~\ref{sec:method_spatially_varying_prior},
 this eventually prevents the prior from adversely affecting the
 reconstruction of new regions in the test. We refer to this method as
 a \textit{spatially-varying} prior-based reconstruction
 routine, that still uses \textit{all} the previous scans in the set $S$.

>>>>

In conventional data acquisition techniques, the tomographic
measurements $\boldsymbol{y}$ are acquired by sampling the physical
object $\boldsymbol{x}$ uniformly with a substantial number of views,
ideally above the Nyquist rate. In such a case when there are
sufficient measurements, reconstruction using the conventional
filtered backprojection (FBP) suffices, as seen in
Fig.~\ref{fig:story} (first column, 600 views). The figure shows the
ground truth image (260$\times$260) of naturally growing sprouts at the top
left (the details of this dataset is postponed to
Section~\ref{sec:results_uniform_prior}).

@@

 \begin{figure}[t]
\centering
	\includegraphics[width=0.45\textwidth]{../images/story/story.png}
        \caption{Illustration of the use of various priors in the
          reconstruction as the number of views reduces. In all cases,
          the ground truth (260$\times$260) appears in the top left.
          For extremely large number of views (600 views), the FBP
          reconstruction (first column) is of very good quality
          (Structural Similarity Index Metric, SSIM=0.92). As the
          number of views become limited (100 views), the
          reconstruction using sparsity prior (middle column) is
          better both visually and quantitatively than FBP (SSIM of
          0.90 vs 0.83). If the number of views is drastically reduced
          (30 views), the presence of both sparsity prior and object
          prior improves the reconstruction (SSIM=0.85) when compared
          to presence of only sparsity prior (SSIM=0.82) or no prior
          (SSIM=0.69).}
 \label{fig:story}
 \end{figure}

 When the number of views is significantly reduced (this paper), the
 sparsity prior alone is not sufficient.  In such cases additional
 information specific to the object being scanned (i.e., object prior)
 is useful in further improving the reconstruction.  While this has
 been done in the context of dynamic CT-scans (as discussed in
 Section~\ref{sec:related}), in this paper we use a set $S$ of
 previous scans of the same object. This is relevant in a longitudinal
 context: the acquisition of sequential CT scans of the same subject
 in order to track time-evolving changes within the subject's
 interior. As seen in Fig.~\ref{fig:story} (last column, 30 views) our
 method uses the prior scans \emph{uniformly} by creating an
 eigenspace representation (Section~\ref{sec:method_uniform_prior}).

 However, with the use of such an object prior, a new challenge
 emerges. The prior set $S$ may potentially overwhelm the necessary
 details; when several prior scans are available, finding the right
 prior is an issue. We therefore seek an algorithm that estimates the
 location and magnitude of new changes in the (unknown) test. As we
 show in Section~\ref{sec:method_spatially_varying_prior}, this
 eventually prevents the prior from adversely affecting the
 reconstruction of new regions in the test. We refer to this method as
 a \textit{spatially-varying} prior-based reconstruction routine, that
 still uses \textit{all} the previous scans in the set $S$, without
 using simply the prior scan, or endeavoring to find the ``right'' prior.
